---
title: "SDG Indicator Review"
#author: "Fiona Spooner"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
    toc: true
geometry: margin=3cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, dpi = 100, fig.align="center")
source("sdg_useful.R")
```

```{r Setup, echo = FALSE}
library(dplyr)
library(gapminder)
library(ggplot2)
library(ggridges)
library(janitor)
library(kableExtra)
library(knitr)
library(leaflet)
library(patchwork)
library(plotly)
library(RColorBrewer)
library(readr)
library(readxl)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(sf)
library(tidyr)
```



```{r, echo = FALSE}
#SDG Tracker Gaps

all_inds_fp <-
  list.files("Data",
             pattern = "*I_",
             recursive = TRUE,
             full.names = TRUE)

full_country_seq <- function(file_path) {
  dat <- clean_names(read_csv(file_path))
  goal <- file_path %>% gsub("Data/", "", .) %>% gsub("/T.*", "", .)
  target <-
    file_path %>% gsub(paste0("Data/", goal, "/"), "", .) %>% gsub("/I_.*", "", .)
  indicator <-
    file_path %>% gsub(paste0("Data/", goal, "/", target, "/"), "", .) %>% sub("_[^_]+$", "", .)
  
  year_out <- dat %>%
    group_by(entity, code) %>%
    summarise(
      latest_data_year = max(year),
      goal = goal,
      target = target,
      indicator = indicator,
      indicator_desc = colnames(.)[4]
    )
  
  return(year_out)
  
}

last_data_df <- lapply(all_inds_fp, full_country_seq)

year_df <- do.call("rbind", last_data_df)
```


```{r, echo = FALSE}
year_df$goal <- year_df$goal %>%
  gsub("Goal_", "", .)

year_df$target <- year_df$target %>%
  gsub("Target_", "", .) %>% gsub("_", ".", .)

year_df$indicator <- year_df$indicator %>%
  gsub("I_", "", .) %>% gsub("_", ".", .)


year_df <- year_df %>% 
  group_by(entity, indicator) %>% 
  filter(latest_data_year == max(latest_data_year))

```

```{r, echo = FALSE, eval = FALSE}
#write.csv(year_df, "sdg_tracker_data_2020_03_09.csv", row.names = FALSE)
```


```{r, echo = FALSE}
#Removing countries which are in the sdg tracker data but do not have a geometry in the #world sf - need to try downloading the rnaturalearthhighres package.

world <- clean_names(ne_countries(scale = "small", returnclass = "sf"))

world$iso_a3[world$admin == "France"] <- "FRA"
world$iso_a3[world$admin == "Indian Ocean Territory"] <- "IOT"
world$iso_a3[world$admin == "Kosovo"] <- "RKS"
world$iso_a3[world$admin == "Norway"] <- "NOR"

year_df <- year_df[year_df$code %in% world$iso_a3 & !is.na(year_df$code),]

```



```{r Loading and Cleaning Indicator Data, echo = FALSE}
inds <- clean_names(read_xlsx("Data/IndicatorSummary.xlsx"))

inds$indicator <-  inds$official_indicator %>% 
  gsub(":.*","",.) %>% 
  gsub("Indicator ", "", .)
```


```{r, echo = FALSE}

track_inds <- inds %>%
  dplyr::select(
    pillar,
    sdg,
    target,
    indicator,
    indicator_tier,
    sdg_tracker_indicator_available,
    sdgt_earliest_date,
    sdgt_date_of_most_recent_data) %>%
  distinct()%>% 
  group_by(indicator) %>% 
  tidyr::fill(pillar, .direction = "updown") %>% 
  ungroup()
```  


```{r, echo = FALSE}
countries <- rep(unique(year_df$entity), each = length(track_inds$indicator))
codes <- rep(unique(year_df$code), each = length(track_inds$indicator))

c_inds <-
  data.frame(countries = I(countries),
             code = codes,
             indicator = I(track_inds$indicator))

all_inds <- merge(c_inds, track_inds, by = "indicator", all = TRUE)

glob_inds <-
  merge(
    year_df,
    all_inds,
    by.x = c("entity", "code","indicator"),
    by.y = c("countries", "code","indicator"),
    all = TRUE
  ) %>%
  select(pillar, indicator_tier, goal, target.x, indicator, entity, code,  latest_data_year,sdg_tracker_indicator_available) %>% 
  arrange(indicator) %>% 
  rename(target = target.x) %>% 
  distinct() %>%
  group_by(indicator) %>% 
  tidyr::fill(pillar, .direction = "updown") %>%
  tidyr::fill(target, .direction = "updown") %>%
  ungroup() 

glob_inds$goal <- gsub("\\..*", "",glob_inds$indicator)

```

# Indicator Availability

There are 244 SDG indicators that are used to measure progress toward the 17 Sustainable Development Goals. However, data availability for these indicators is varied with some indicators having better data coverage than others

We downloaded all the indicator data available on [SDG Tracker](https://sdg-tracker.org/). This plot shows the proportion of indicators with data available for each of the SDGs


```{r, echo = FALSE}

ind_ord <- unique(glob_inds$indicator)

glob_inds <- glob_inds %>%
  mutate(got_indicator = ifelse(grepl("Yes", sdg_tracker_indicator_available), 1, 0))

glob_inds$indicator <- factor(glob_inds$indicator, levels = ind_ord)

#write.csv(glob_inds, "global_indicators.csv", row.names = FALSE)

```


```{r, echo = FALSE, fig.height= 7, fig.width= 9}

sdg_plot <- glob_inds %>% 
  dplyr::select(goal,indicator, got_indicator) %>% 
  group_by(goal) %>% 
  distinct() %>% 
  mutate(sum_inds = sum(got_indicator), num_inds = n(), prop_inds = (sum_inds/num_inds)) %>% 
  select(goal, sum_inds, num_inds, prop_inds) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(goal = factor(goal, levels = 1:17)) %>% 
  ggplot(., aes(x = goal, y = prop_inds, group = goal,fill = goal, colour = goal))+
  geom_bar(stat = "identity", colour = "black") +
  sdg_fill_scale + 
  sdg_col_scale + 
  theme_fiona()+
  theme(legend.position = "none") + 
  ylab("Proportion of Indicator\n Available") + 
  xlab("Sustainable Development Goal") + 
  ggtitle("Indicators Available on SDG Tracker") + 
  ylim(0,1)

sdg_plot

```

### Indicator tiers across goals

Each of the SDG indicators have been classified into three tiers by the [Inter-agency and Expert Group on SDG Indicators](https://unstats.un.org/sdgs/iaeg-sdgs/). 

  * Tier 1: Indicator is conceptually clear, has an internationally established methodology and standards are available, and data are regularly produced by countries for at least 50 per cent of countries and of the population in every region where the indicator is relevant.
  
  * Tier 2: Indicator is conceptually clear, has an internationally established methodology and standards are available, but data are not regularly produced by countries.
  
  * Tier 3: No internationally established methodology or standards are yet available for the indicator, but methodology/standards are being (or will be) developed or tested.


Here we look at how the distribution of the tiers across the SDGs. 
```{r, echo = FALSE, fig.height= 7, fig.width= 9}
tier_goal_plot <- glob_inds %>% 
  dplyr::select(goal,indicator_tier,indicator) %>% 
  group_by(goal) %>% 
  distinct() %>% 
  ungroup() %>% 
  arrange(indicator_tier) %>% 
  mutate(goal = factor(goal, levels = 1:17)) %>% 
  ggplot(., aes(x = goal,fill = indicator_tier))+
  geom_bar(position = "fill", colour = "black") +
  tier_fill_scale +
  theme_fiona()+
  ylab("Indicator Tier Breakdown") + 
  xlab("Sustainable Development Goal") + 
  ggtitle("Indicator Tiers Across Goals") + 
  guides(fill=guide_legend(title="Indicator Tier")) +
  ylim(0,1) 
 

tier_goal_plot

```

### Indicator tiers available on SDG Tracker


The SDGs can be grouped into three broad pillars:
  
  * Biosphere (SDGs 6, 7, 12, 13, 14, 15)
  * Economy (SDGs 1, 2, 3, 8, 9)
  * Society (SDGs 4, 5, 10, 11, 16, 17)

Here we look at the indicator data availability across these pillars based on the SDG Tracker data.

```{r, echo = FALSE, fig.height= 7, fig.width= 9}
pillar_plot <- glob_inds %>% 
  dplyr::select(pillar,indicator, got_indicator) %>% 
  group_by(pillar) %>% 
  distinct() %>% 
  mutate(sum_inds = sum(got_indicator), num_inds = n(), prop_inds = (sum_inds/num_inds)) %>% 
  select(pillar, sum_inds, num_inds, prop_inds) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(pillar = factor(pillar, levels = c("Biosphere", "Society", "Economy"))) %>% 
  ggplot(., aes(x = pillar, y = prop_inds, group = pillar,fill = pillar, colour = pillar))+
  geom_bar(stat = "identity", colour = "black") +
  pillar_fill_scale + 
  pillar_col_scale + 
  theme_fiona()+
  theme(legend.position = "none") + 
  ylab("Proportion of \nIndicators Available") + 
  xlab("Pillar") + 
  ggtitle("Indicators Available on SDG Tracker") +
  ylim(0,1)

pillar_plot


```

### Indicator tiers across SDG pillars

We also look at the distribution of the indicator tiers across the SDG pillars.


```{r, echo = FALSE, fig.height= 7, fig.width= 9}
pillar_tier_plot <- glob_inds %>% 
  dplyr::select(pillar,indicator_tier,indicator) %>% 
  group_by(pillar) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(pillar = factor(pillar, levels = c("Biosphere", "Society", "Economy"))) %>% 
  ggplot(., aes(x = pillar,fill = indicator_tier))+
  geom_bar(position = "fill", colour = "black") +
  theme_fiona()+
  tier_fill_scale +
  ylab("Proportion of Pillar Made\n up of Each Indicator Tier") + 
  xlab("Pillar") + 
  ggtitle("Indicator Tiers Across SDG Pillars") +
  guides(fill=guide_legend(title="Indicator Tier")) 

pillar_tier_plot


```

```{r, echo = FALSE}
kable(
  inds %>%
    group_by(pillar) %>%
    summarise(
      count = n(),
      no_track_count = sum(sdg_tracker_indicator_available == "No"),
      tier_3 = sum(indicator_tier == 3),
      prop_no_ind = round((no_track_count / count) * 100, 2),
      prop_t3 = round((tier_3 / count) * 100, 2)
    ) %>%
    select(pillar, prop_no_ind, prop_t3) %>%
    arrange(-prop_no_ind),
  col.names = c("Pillar",
                "Missing Indicators (%)",
                "Tier 3 Indicators (%)")
) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Indicator Availability

### The number of countries which have data for an indicator - at goal level

There are a number of indicators which are only available for a limited number of countries. Here we highlight the number of countries with data available for each indicator.

```{r, echo = FALSE, fig.height = 10, fig.width = 8}
num_countries <- glob_inds %>%
  select(goal, code, indicator, entity, latest_data_year) %>%
 # group_by(goal, indicator, entity) %>%
  filter(code != "ATA") %>%
  #ungroup() %>%
  group_by(goal, indicator) %>%
  summarise(num_countries = sum(!is.na(latest_data_year))) %>%
  ungroup() %>%
  mutate(goal = factor(goal, levels = 17:1)) %>%
  arrange(num_countries) %>%
  ggplot(., aes(x = num_countries, y = goal, fill = goal)) +
  geom_density_ridges(
    scale = 1.5,
    gradient_lwd = 1.,
    alpha = 0.6,
    from = 0,
    to = 233,
    jittered_points = TRUE,
    position = position_points_jitter(width = 0.075, height = 0),
    point_shape = '|',
    point_size = 1.5,
    point_alpha = 1,
    alpha = 0.7
  ) +
  theme_fiona() +
  sdg_fill_scale +
  theme(legend.position = "none") +
  ylab("Sustainable Development Goal") +
  xlab("Number of Countries with Indicator Data")

num_countries


```

### The number of indicators available in each country:

 <style>
 .html-widget {
     margin: auto;
 }
 </style>


```{r, echo = FALSE, fig.height= 6, fig.width= 9}

world_sf <- world %>%
  select(iso_a3, region_un, continent, pop_est, gdp_md_est) %>%
  mutate(code = iso_a3, region_un = as.factor(region_un)) %>%
   left_join(., glob_inds, by = "code") #%>% 
  # select(iso_a3, region_un, code, pillar, indicator_tier, indicator,entity, latest_data_year, got_indicator) %>% st_write(.,  "global_indicators.shp")

num_ind_countries <- world_sf %>%
  group_by(entity, indicator) %>%
  filter(
    latest_data_year == max(latest_data_year) |
      is.na(latest_data_year) &
      region_un != "Antarctica" &
      region_un != "Seven seas (open ocean)") %>%
  ungroup() %>%
  group_by(entity) %>%
  summarise(sum_ind = sum(!is.na(latest_data_year)))

num_ind_countries <- as_Spatial(num_ind_countries)

bins <- c(0, 25, 50, 75, 100, 125, 150, 175, 200, 225, 244)
pal <-
  colorBin("YlOrRd", domain = num_ind_countries$sum_ind, bins = bins)

labels <- sprintf(
  "<strong>%s</strong><br/>%g indicators",
  num_ind_countries$entity,
  num_ind_countries$sum_ind
) %>% lapply(htmltools::HTML)

leaflet(num_ind_countries) %>%
  setView(0, 30, 2) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    fillColor = ~ pal(sum_ind),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE
    ),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto"
    )
  ) %>%
  addLegend(
    pal = pal,
    values = ~ sum_ind,
    opacity = 0.7,
    title = NULL,
    position = "bottomleft",
      labFormat = labelFormat(big.mark = "")
  )

```

```{r, echo = FALSE, eval = FALSE}
#saveRDS(world_sf, "global_indicators.rds")

```

```{r, eval = FALSE, echo = FALSE}
### Table of indicators with covering the fewest countries - <30 

inds_min <- inds %>%
  select(official_indicator, indicator)

kable(
  glob_inds %>%
    # group_by(entity, indicator) %>%
    # filter(
    #   latest_data_year == max(latest_data_year) |
    #     is.na(latest_data_year)
    # ) %>%
    # ungroup() %>%
    group_by(indicator) %>%
    summarise(sum_no_country = sum(is.na(latest_data_year))) %>%
    arrange(-sum_no_country) %>%
    filter(sum_no_country > 205) %>%
    select(indicator, sum_no_country) #%>% 
#   left_join(., inds_min, by = "indicator"),
#   col.names = c("Indicator", "Indicator Description"),
#   caption = "Indicators with the oldest data"
# ) 

```

# Age of Indicator Data

### The age of the most recent indicator data - all countries 

The SDG Tracker data has country level data and records a time-series of the data available for each country, in each indicator. Here we show the year with the most recent data available for each country, across the SDGs.

```{r, echo = FALSE, fig.height= 7, fig.width= 9}

goal_ind_age <- glob_inds %>% 
  filter(!is.na(latest_data_year)) %>% 
  # group_by(goal,indicator, entity) %>% 
  # filter(latest_data_year == max(latest_data_year)) %>% 
  # ungroup() %>% 
  group_by(goal, latest_data_year) %>% 
  mutate(year_freq = n()) %>%
  ungroup() %>% 
  mutate(goal = factor(goal, levels = 17:1)) %>% 
  select(goal, latest_data_year, year_freq) %>% 
  arrange(goal,latest_data_year) %>% 
  distinct() %>% 
  ggplot(., aes(x = latest_data_year, y = goal, size = year_freq, fill = goal, col = goal, text = year_freq))+
  geom_point(alpha = 0.75, shape=21)+ 
  sdg_fill_scale + 
  sdg_col_scale +
  theme_fiona() + 
  theme(legend.position = "none") + 
  ylab("Sustainable Development Goal") + 
  xlab("Year") + 
  ggtitle("Age of Most Recent Indicator Data - All Countries")+
  guides(fill = FALSE, col = FALSE, size=guide_legend(title="Number of Indicators")) 


fig <- ggplotly(goal_ind_age, tooltip = "text")
 
fig

```



### The age of the most recent indicator data by region

This plot shows the same data but broken down by region.

```{r, echo = FALSE, fig.height= 7, fig.width= 9}

world_df <- data.frame(world %>%
  select(iso_a3, region_un, continent, pop_est, gdp_md_est) %>%
  mutate(code = iso_a3, region_un = as.factor(region_un)) %>%
  left_join(., glob_inds, by = "code"))

goal_ind_age_region <- world_df %>% 
  filter(!is.na(latest_data_year)& region_un != "Antarctica"& region_un != "Seven seas (open ocean)") %>% 
  group_by(goal,indicator, entity) %>% 
  filter(latest_data_year == max(latest_data_year)) %>% 
  ungroup() %>% 
  group_by(region_un ,goal, latest_data_year) %>% 
  mutate(year_freq = n()) %>%
  ungroup() %>% 
  mutate(goal = factor(goal, levels = 17:1)) %>% 
  select(region_un, goal, latest_data_year, year_freq) %>% 
  arrange(goal,latest_data_year) %>% 
  distinct() %>% 
  ggplot(., aes(x = latest_data_year, y = goal, size = year_freq, fill = goal, col = goal, text = year_freq), group = region_un)+
  geom_point(alpha = 0.75, shape=21)+ 
  sdg_fill_scale + 
  sdg_col_scale +
  theme_fiona() + 
  theme(legend.position = "none") + 
  ylab("Sustainable Development Goal") + 
  xlab("Year") + 
  ggtitle("Age of Most Recent Indicator Data - By Region")+
  facet_wrap( ~ region_un) + 
  guides(fill = FALSE, col = FALSE, size=guide_legend(title="Number of Indicators"))+
  theme(panel.spacing = unit(2, "lines"))


#goal_ind_age_region 

fig <- ggplotly(goal_ind_age_region, tooltip = "text")
 
fig


```

### Table of indicators with the oldest data (pre 2000)

```{r, echo = FALSE}

inds_min <- inds %>% 
  select(official_indicator, indicator)

kable(world_df %>% 
  filter(!is.na(latest_data_year)& region_un != "Antarctica"& region_un != "Seven seas (open ocean)") %>% 
  group_by(indicator) %>% 
    #filter(latest_data_year == min(latest_data_year)) %>% 
    mutate(min_year = min(latest_data_year), max_year = max(latest_data_year)) %>% 
    arrange(min_year) %>% 
    select(indicator, min_year, max_year) %>% 
       distinct() %>% 
    filter(min_year <= 2000) %>% 
    left_join(.,inds_min, by = "indicator") %>% 
    select(indicator, official_indicator, min_year, max_year), col.names = c("Indicator", "Indicator Description", "Oldest Data", "Newest Data"), caption = "") 


```

 <style>
 .html-widget {
     margin: auto;
 }
 </style>


### The average date of the most recent available indicator data - at the country level

```{r, echo = FALSE, fig.height= 6, fig.width= 9}


mean_yr_countries <- world_sf %>%
  group_by(entity) %>%
  filter(!is.na(latest_data_year)) %>% 
  summarise(mean_year = mean(latest_data_year))


mean_yr_countries <- as_Spatial(mean_yr_countries)

bins <- c(2012:2020)
pal <-
  colorBin("YlOrRd", domain =mean_yr_countries$mean_year, bins = bins)

labels <- sprintf(
  "<strong>%s</strong><br/>Average year of data availability: %g",
  mean_yr_countries$entity,
  mean_yr_countries$mean_year
) %>% lapply(htmltools::HTML)

leaflet(mean_yr_countries) %>%
  setView(0, 30, 2) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    fillColor = ~ pal(mean_year),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE
    ),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto"
    )
  ) %>%
  addLegend(
    pal = pal,
    values = ~ mean_year,
    opacity = 0.7,
    title = NULL,
    position = "bottomleft",
      labFormat = labelFormat(big.mark = "")
  )

```

### Table of countries with the oldest data

```{r, echo = FALSE}

inds_min <- inds %>% 
  select(official_indicator, indicator)

kable(world_df %>% 
  filter(!is.na(latest_data_year)& region_un != "Antarctica"& region_un != "Seven seas (open ocean)") %>% 
  select(entity,indicator, latest_data_year) %>% 
  group_by(entity) %>% 
 #filter(latest_data_year == min(latest_data_year)) %>% 
    mutate(min_year = min(latest_data_year), max_year = max(latest_data_year)) %>% 
   ungroup() %>% 
   arrange(min_year) %>% 
    select(entity, min_year, max_year, indicator, latest_data_year) %>% 
   filter(min_year == latest_data_year, min_year <= 2000) %>% 
       distinct() %>% 
   left_join(., inds_min, by = "indicator") %>% 
  select(entity, min_year, indicator, official_indicator,max_year) 
  , caption = "", col.names = c("Country","Country Oldest Data", "Indicator w/ Oldest Data","Indicator Description","Country Newest Data")) 


```



```{r, eval = FALSE, echo = FALSE}

ggplot(glob_inds, aes(x = indicator, y = latest_data_year, group = pillar, colour = pillar))+
  geom_point()+
  theme_bw()


glob_inds %>% 
  group_by(entity) %>% 
  summarise(average_data_year = mean(na
                                     .omit(latest_data_year))) %>% 
  arrange(average_data_year)
  

```



```{r, eval = FALSE, echo = FALSE}

#The range of most recent dates for each indicator

date_ranges <- glob_inds %>% 
  group_by(indicator) %>% 
  summarise(earliest_country_data = min(na.omit(latest_data_year)), latest_country_data = max(na.omit(latest_data_year))) 

date_ranges$earliest_country_data[is.infinite(date_ranges$earliest_country_data)] <- NA

date_ranges$latest_country_data[is.infinite(date_ranges$latest_country_data)] <- NA

```



```{r, echo = FALSE, eval = FALSE}

#Counting the number of indicators without data in each country

ind_count <- glob_inds %>%
  filter(!is.na(code)) %>%
  group_by(entity, code) %>%
  summarise(got_ind = sum(!is.na(latest_data_year)), missing_ind = 244 - got_ind) %>% 
  ungroup() %>% 
  #arrange(-got_ind) %>% 
  mutate(rank = rank(missing_ind))

```



```{r, echo = FALSE, eval = FALSE}
kable(ind_count %>% 
  dplyr::select(entity, got_ind) %>% 
  arrange(got_ind) %>% 
  head(., 20))
```


```{r, eval = FALSE, echo = FALSE}


#Exploring the how GDP and population impact the indicator coverage of countries:

#Number of indicators per country vs GDP


```



```{r, echo = FALSE, eval = FALSE}
inds %>% 
  group_by(indicator_tier, pillar) %>% 
#  summarise(count = n(),no_track_count = sum(sdg_tracker_indicator_available == "No"), prop_no = no_track_count/count)
  filter(indicator_tier == 3) %>% 
  select(official_indicator, sdgt_date_of_most_recent_data) %>% 
  distinct()

kable(inds$sdgt_data_publisher)

```

```{r, echo = FALSE, eval = FALSE}

jit_tier_track <- ggplot(glob_inds, aes(x = got_indicator, y = indicator_tier))+
  geom_jitter()
  
  
  

```
